---
title: "K-nearest neighbour Classification Modelling"
output: html_document
runtime: shiny
---

Note :  1. This application works only for 2 or more predictors(independent variables)
        2. The general rule of thumb of square root of number of rows in validation data is folllowed in figuring out the ideal value of k

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(caret)
library(e1071)
library(rhandsontable)
library(datasets)
library(psycho)
#library(lattice)
library(FNN)

```

```{r,echo=FALSE}
sidebarPanel(
  checkboxInput("ex","Uncheck for using your own file",value = TRUE),
  fileInput("file", "Upload the *.csv file with headers"),
  selectInput("dt","Choose Train/Valid/Full for Visuals",choices = c("Train","Valid","Full"),selected = "Full"),
  sliderInput("train_num", label = "Enter the proportion of training dataset:",
                     min = 0.6, max = 1, value = 0.6, step = 0.01),
  checkboxInput("std","Standardize",value = FALSE),
  
  uiOutput("vx"),
  uiOutput("vy"),
  downloadButton("downloadPlot", "Download Plot") 
  
)
mainPanel(
   tabsetPanel(type = "tab",
                tabPanel("Model Summary", verbatimTextOutput("AD") ),
                tabPanel("Model Visualization", plotOutput("MV") ),
                tabPanel("Model Evaluation",verbatimTextOutput("ME")),
                tabPanel("Model Deployment",verbatimTextOutput("MD"))
                ),
    h6("Edit the test data record"),
   rHandsontableOutput("testdata"),
   h6("", tags$img(src ="K.JPG", height= 400, width=400))
)
output$AD<-renderPrint({
   if(input$ex == TRUE)
     {data("iris")
      data = iris
    gp = runif(nrow(data))
    data[,]= data[order(gp),]
      }
     else{
  file1 = input$file
     if(is.null(file1)){return()}
     
     data =  read.table(file =  file1$datapath,sep =",",header = TRUE)
     if(is.null(data())){return()}
     }
     ds = data
     
     ds = select(ds,input$variablex)
     if(ncol(ds)==2){return()}
     if(input$std == TRUE)
       {ds = standardize(ds)}
     options(scipen = 999)
       prop = input$train_num
       set.seed(1)
       dataframe =  ds
       train.rows = sample(row.names(dataframe),dim(dataframe)[1]*prop)
       dataframet = dataframe[train.rows,]
       valid.rows = setdiff(row.names(dataframe),train.rows)
       dataframev = dataframe[valid.rows,]    
 
      
       indexdependent= grep(input$variabley, colnames(dataframe))
      # print(indexdependent)
     if(prop<1)
     {  accuracy.df = data.frame(k =seq(1,sqrt(nrow(dataframev)),accuracy = rep(0,sqrt(nrow(dataframev)))))
       
       
      for(i in 1:sqrt(nrow(dataframev)) )
{
  prediction= knn(train =dataframet[,-indexdependent],test = dataframev[,-indexdependent],cl =  dataframet[,indexdependent],k= i)
       accuracy.df[i,2]=confusionMatrix(prediction,dataframev[,indexdependent])$overall[1]
      }   
       names(accuracy.df)=c("k", "Accuracy")
       accuracy.df[,] = accuracy.df[order(1-accuracy.df$Accuracy),]
     print(accuracy.df)  
     cat(sprintf("The value for k to be chosen is %d ",accuracy.df[1,1]))
   # print("The ideal value for the ")
     
     }
     
       else 
       {
         print("There needs to be a validation/test data set")
       }
      

     
     
})
output$MV<-renderPlot({ 
  if(input$ex == TRUE)
     {data("iris")
      data = iris
    gp = runif(nrow(data))
    data[,]= data[order(gp),]
      }
     else{
  file1 = input$file
     if(is.null(file1)){return()}
     
     data =  read.table(file =  file1$datapath,sep =",",header = TRUE)
     if(is.null(data())){return()}
     }
     ds = data
     
     ds = select(ds,input$variablex)
     if(ncol(ds)==2){return()}
     if(input$std == TRUE)
       {ds = standardize(ds)}
     dataframe = ds
  
     indexdependent= grep(input$variabley, colnames(dataframe))
     prop = input$train_num
       set.seed(1)
       dataframe =  ds
       train.rows = sample(row.names(dataframe),dim(dataframe)[1]*prop)
       dataframet = dataframe[train.rows,]
       valid.rows = setdiff(row.names(dataframe),train.rows)
       dataframev = dataframe[valid.rows,]    
 
   #  super.sym <- trellis.par.get("superpose.symbol")
   #  ngp =  nlevels(dataframe[,indexdependent])
 # splom(~dataframe[,-indexdependent]|dataframe[,indexdependent], data=dataframe) 
  # parallelplot(~dataframe[,-indexdependent] | dataframe[,indexdependent], dataframe)           
       if(input$dt == "Full")   
   {
         if(ncol(dataframe)>2)
       {pairs(dataframe[,-indexdependent], col = dataframe[,indexdependent], oma=c(3,3,3,15))
par(xpd = TRUE)
legend("bottomright", fill = unique( dataframe[,indexdependent]), legend = c( levels( dataframe[,indexdependent])))
         }
         
  }
       if(input$dt == "Train")   
   { 
        if(ncol(dataframet)>2)
       {  pairs(dataframet[,-indexdependent], col = dataframet[,indexdependent], oma=c(3,3,3,15))
par(xpd = TRUE)
legend("bottomright", fill = unique( dataframet[,indexdependent]), legend = c( levels( dataframet[,indexdependent])))
        }
         
  }
       
         if(input$dt == "Valid")   
   { 
           if(ncol(dataframev)>2 && prop<1)
         { pairs(dataframev[,-indexdependent], col = dataframev[,indexdependent], oma=c(3,3,3,15))
par(xpd = TRUE)
legend("bottomright", fill = unique( dataframev[,indexdependent]), legend = c( levels( dataframev[,indexdependent]))) 
           }
   }
       
       
     
})

output$ME<-renderPrint({
   if(input$ex == TRUE)
     {data("iris")
      data = iris
    gp = runif(nrow(data))
    data[,]= data[order(gp),]
      }
     else{
  file1 = input$file
     if(is.null(file1)){return()}
     
     data =  read.table(file =  file1$datapath,sep =",",header = TRUE)
     if(is.null(data())){return()}
     }
     ds = data
     
     ds = select(ds,input$variablex)
     if(ncol(ds)==2){return()}
     if(input$std == TRUE)
       {ds = standardize(ds)}
     mod = paste(input$variabley,"~.")
     options(scipen = 999)
       prop = input$train_num
       set.seed(1)
       dataframe =  ds
       train.rows = sample(row.names(dataframe),dim(dataframe)[1]*prop)
       dataframet = dataframe[train.rows,]
       valid.rows = setdiff(row.names(dataframe),train.rows)
       dataframev = dataframe[valid.rows,]    
 
      
       indexdependent= grep(input$variabley, colnames(dataframe))
     #  print(indexdependent)
     if(prop<1)
     {  accuracy.df = data.frame(k =seq(1,sqrt(nrow(dataframev)),accuracy = rep(0,sqrt(nrow(dataframev)))))
       
       
      for(i in 1:sqrt(nrow(dataframev)) )
{
  prediction= knn(train =dataframet[,-indexdependent],test = dataframev[,-indexdependent],cl =  dataframet[,indexdependent],k= i)
       accuracy.df[i,2]=confusionMatrix(prediction,dataframev[,indexdependent])$overall[1]
      }   
       names(accuracy.df)=c("k", "Accuracy")
       accuracy.df[,] = accuracy.df[order(1-accuracy.df$Accuracy),]
   #  print(accuracy.df)  
     #cat(sprintf("The value for k to be chosen is %d ",accuracy.df[1,1]))
   # print("The ideal value for the ")
       ki = accuracy.df[1,1]
  prediction= knn(train =dataframet[,-indexdependent],test = dataframev[,-indexdependent],cl =  dataframet[,indexdependent],k= ki)
  print(confusionMatrix(prediction,dataframev[,indexdependent]))
     
     }
     
       else 
       {
         print("There needs to be a validation/test data set")
       }
      

   

     
     
})

output$MD<-renderPrint({
  if(input$ex == TRUE)
     {data("iris")
      data = iris
    gp = runif(nrow(data))
    data[,]= data[order(gp),]
      }
     else
     {
  file1 = input$file
     if(is.null(file1)){return()}
     
     data =  read.table(file =  file1$datapath,sep =",",header = TRUE)
     if(is.null(data())){return()}
     }
     ds = data
     
     ds = select(ds,input$variablex)
     if(ncol(ds)==2){return()}
     if(input$std == TRUE)
       {ds = standardize(ds)}
    
     options(scipen = 999)
       prop = input$train_num
       set.seed(1)
       dataframe =  ds
       train.rows = sample(row.names(dataframe),dim(dataframe)[1]*prop)
       dataframet = dataframe[train.rows,]
       valid.rows = setdiff(row.names(dataframe),train.rows)
       dataframev = dataframe[valid.rows,]  
       
       indexdependent= grep(input$variabley, colnames(dataframe))
       if(prop<1)
     {  accuracy.df = data.frame(k =seq(1,sqrt(nrow(dataframev)),accuracy = rep(0,sqrt(nrow(dataframev)))))
       
       
      for(i in 1:sqrt(nrow(dataframev)) )
{
  prediction= knn(train =dataframet[,-indexdependent],test = dataframev[,-indexdependent],cl =  dataframet[,indexdependent],k= i)
       accuracy.df[i,2]=confusionMatrix(prediction,dataframev[,indexdependent])$overall[1]
      }   
       names(accuracy.df)=c("k", "Accuracy")
       accuracy.df[,] = accuracy.df[order(1-accuracy.df$Accuracy),]
    # print(accuracy.df)  
   # cat(sprintf("The value for k to be chosen is %d ",accuracy.df[1,1]))
   # print("The ideal value for the ")
       ki = accuracy.df[1,1]
     
     }
     
       else 
       {
         print("There needs to be a validation/test data set")
       }
      
# if(input$mt == "lda")
    # {
      # model = MASS::lda(formula = as.formula(mod),data = dataframet)
       test_data = data.frame(hot_to_r(input$testdata))
       if(ncol(test_data)== 1)
       { df  = select(dataframet,-c(input$variabley))
         colnames(test_data)=  colnames(df)
       }
       #print(dim(dataframet))
      # print(dim(test_data))
       if(ncol(test_data)>1 && prop <1)
     {  prediction= knn(train =dataframet[,-indexdependent],test= test_data,cl =  dataframet[,indexdependent],k= ki)  
     test_data$predictedvalue = prediction
       print(test_data)
       
     } 
   

     
     
})

output$vx <- renderUI({
   
 if(input$ex == TRUE)
     {data("iris")
      data = iris}
     else
     {
     
     file1 = input$file
     if(is.null(file1)){return()}
     
     data =  read.table(file =  file1$datapath,sep =",",header = TRUE)
     if(is.null(data())){return()}
    
     }
     
    checkboxGroupInput("variablex","Select the variables",choices = colnames(data),selected = colnames(data))
    
    
  


  })
output$vy <- renderUI({
   
 if(input$ex == TRUE)
     {data("iris")
      data = iris}
     else
     {
     
     file1 = input$file
     if(is.null(file1)){return()}
     
     data =  read.table(file =  file1$datapath,sep =",",header = TRUE)
     if(is.null(data())){return()}
     }
    ds = data
    ds = select(ds,input$variablex) 
    ds = select_if(ds,is.factor)
     
    selectInput("variabley","Select the dependent variable",choices = colnames(ds),selected = "" )


  })
output$k <- renderUI({
   
 if(input$ex == TRUE)
     {data("iris")
      data = iris}
     else
     {
     
     file1 = input$file
     if(is.null(file1)){return()}
     
     data =  read.table(file =  file1$datapath,sep =",",header = TRUE)
     if(is.null(data())){return()}
     }
    ds = data
    prop = input$train_num
       set.seed(1)
       dataframe =  ds
       train.rows = sample(row.names(dataframe),dim(dataframe)[1]*prop)
       dataframet = dataframe[train.rows,]
       valid.rows = setdiff(row.names(dataframe),train.rows)
       dataframev = dataframe[valid.rows,] 
   
    sliderInput("k", label = "Enter the value of k ",
                     min = 1, max = NROW(dataframet), value = 1, step = 1)


  })

output$testdata <- renderRHandsontable({
   
 if(input$ex == TRUE)
     {data("iris")
      data = iris}
     else{
     
     file1 = input$file
     if(is.null(file1)){return()}
     
     data =  read.table(file =  file1$datapath,sep =",",header = TRUE)
     if(is.null(data())){return()}
     }
    
    ds = data
    ds = select(ds,input$variablex) 
    if(input$std == TRUE)
       {ds = standardize(ds)}
     ds = select(ds,-c(input$variabley))
    
 
   rhandsontable(data.frame(ds[1,]))
    
   

  })


output$downloadPlot<- downloadHandler(
    filename = function() {
      paste("Scatterplot", ".png", sep = "")
    },
    content = function(file) {
      png(file)
if(input$ex == TRUE)
     {data("iris")
      data = iris
    gp = runif(nrow(data))
    data[,]= data[order(gp),]
      }
     else{
  file1 = input$file
     if(is.null(file1)){return()}
     
     data =  read.table(file =  file1$datapath,sep =",",header = TRUE)
     if(is.null(data())){return()}
     }
     ds = data
     
     ds = select(ds,input$variablex)
     if(ncol(ds)==2){return()}
     if(input$std == TRUE)
       {ds = standardize(ds)}
     dataframe = ds
  
     indexdependent= grep(input$variabley, colnames(dataframe))
     prop = input$train_num
       set.seed(1)
       dataframe =  ds
       train.rows = sample(row.names(dataframe),dim(dataframe)[1]*prop)
       dataframet = dataframe[train.rows,]
       valid.rows = setdiff(row.names(dataframe),train.rows)
       dataframev = dataframe[valid.rows,]    
 
   #  super.sym <- trellis.par.get("superpose.symbol")
   #  ngp =  nlevels(dataframe[,indexdependent])
 # splom(~dataframe[,-indexdependent]|dataframe[,indexdependent], data=dataframe) 
  # parallelplot(~dataframe[,-indexdependent] | dataframe[,indexdependent], dataframe)           
       if(input$dt == "Full")   
   {
         if(ncol(dataframe)>2)
       {pairs(dataframe[,-indexdependent], col = dataframe[,indexdependent], oma=c(3,3,3,15))
par(xpd = TRUE)
legend("bottomright", fill = unique( dataframe[,indexdependent]), legend = c( levels( dataframe[,indexdependent])))
         }
         
  }
       if(input$dt == "Train")   
   { 
        if(ncol(dataframet)>2)
       {  pairs(dataframet[,-indexdependent], col = dataframet[,indexdependent], oma=c(3,3,3,15))
par(xpd = TRUE)
legend("bottomright", fill = unique( dataframet[,indexdependent]), legend = c( levels( dataframet[,indexdependent])))
        }
         
  }
       
         if(input$dt == "Valid")   
   { 
           if(ncol(dataframev)>2 && prop<1)
         { pairs(dataframev[,-indexdependent], col = dataframev[,indexdependent], oma=c(3,3,3,15))
par(xpd = TRUE)
legend("bottomright", fill = unique( dataframev[,indexdependent]), legend = c( levels( dataframev[,indexdependent]))) 
           }
   }
       
       
     
      dev.off()
    })